---
title: "LCA_PE"
author: "Taylor"
date: "2022-11-23"
output: html_document
---

```{r, include=FALSE}

library(haven) ## Reading data - importing datasets
library(tidyLPA) ## Run LPA
library(ggplot2) ## Plotting
library(tidyverse) ## Enables tidy principles
library(mice) ## Missing data and imputations
library(naniar) ## MCAR test

data <- read_sav("1st_wave_collated.sav") 
setwd("C:/Users/taylo/Documents/PE_LCA")

```


## Missing data analysis
```{r echo=FALSE, warning=FALSE, message=FALSE}
c_data <- data %>%
  select(Proteus_1, Proteus_2, Proteus_3, Proteus_3, Proteus_4, Proteus_5, Proteus_6, 198:203)
      

## This calculates % of missing data
Ctotalcells <- prod(dim(c_data))
Cmissingcells <- sum(is.na(c_data))
(Cmissingcells * 100)/(Ctotalcells)

## Testing if missing values are completely at random - MCAR
mcar_test(c_data)

## Highest missing values per variable
c_data %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "names", values_to = "values") %>%
  arrange(desc(values)) %>%
  mutate(perc=(values)/571) %>% # note that 968 is the number of participants and you should replace it with your N 
  slice_head(n=10)
```


## Estimating LPA models (1 to 6) with different number of classes (1:5) 
Here we look for the lowest AIC, BIC, etc. and we follow suggestions outlined in output
```{r echo=FALSE, warning=FALSE, message=FALSE}
set.seed(123)
Models <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(1:5, variances = c("equal", "varying", "equal", "varying"),
                    covariances = c("zero", "zero", "equal", "varying"))%>%
  compare_solutions(statistics = c("AIC","BIC","AWE", "CLC", "KIC"))

Models
```


## Estimating best models 

Model 1 is Equal variances and covariances fixed to 0 (CIDP) 
Model 2 is Varying variances and covariances fixed to 0 (CVDP) 
Model 3 is Equal variances and equal covariances (CIRP) 
Model 4 and 5 are not able to fit Mclust 
Model 6 is Varying variances and Varying covariances (CVUP) 


```{r, echo=FALSE}
## Here we run models of interest to create a plot with BIC/AIC and entropy
set.seed(124)
CIRP_2 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(2, variances="equal",covariances="equal")

set.seed(125)
CIRP_3 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(3, variances="equal",covariances="equal")

set.seed(126)
CIRP_4 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(4, variances="equal",covariances="equal")

set.seed(127)
CIRP_5 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(5, variances="equal",covariances="equal")

set.seed(128)
CIDP_2 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(2, variances="equal",covariances="zero")

set.seed(129)
CIDP_3 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(3, variances="equal",covariances="zero")

set.seed(130)
CIDP_4 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(4, variances="equal",covariances="zero")

set.seed(131)
CIDP_5 <- c_data%>%
  select("Proteus_1", "Proteus_2", "Proteus_3", "Proteus_4", "Proteus_5", "Proteus_6")%>%
  single_imputation()%>%
  estimate_profiles(5, variances="equal",covariances="zero")

fit <- as_tibble(rbind(CIRP_2[["model_3_class_2"]][["fit"]],CIRP_3[["model_3_class_3"]][["fit"]],
      CIRP_4[["model_3_class_4"]][["fit"]],CIRP_5[["model_3_class_5"]][["fit"]],
      CIDP_2[["model_1_class_2"]][["fit"]],CIDP_3[["model_1_class_3"]][["fit"]],
      CIDP_4[["model_1_class_4"]][["fit"]],CIDP_5[["model_1_class_5"]][["fit"]])) %>%
  select(Model,Classes,LogLik,AIC,BIC,AWE,Entropy,n_min,BLRT_p)

## select paramters to plot
aic22 <- as.numeric(fit %>% select(AIC)%>% slice(1))
aic23 <- as.numeric(fit %>% select(AIC)%>% slice(2))
aic24 <- as.numeric(fit %>% select(AIC)%>% slice(3))
aic25 <- as.numeric(fit %>% select(AIC)%>% slice(4))
aic62 <- as.numeric(fit %>% select(AIC)%>% slice(5))
aic63 <- as.numeric(fit %>% select(AIC)%>% slice(6))
aic64 <- as.numeric(fit %>% select(AIC)%>% slice(7))
aic65 <- as.numeric(fit %>% select(AIC)%>% slice(8))
bic22 <- as.numeric(fit %>% select(BIC)%>% slice(1))
bic23 <- as.numeric(fit %>% select(BIC)%>% slice(2))
bic24 <- as.numeric(fit %>% select(BIC)%>% slice(3))
bic25 <- as.numeric(fit %>% select(BIC)%>% slice(4))
bic62 <- as.numeric(fit %>% select(BIC)%>% slice(5))
bic63 <- as.numeric(fit %>% select(BIC)%>% slice(6))
bic64 <- as.numeric(fit %>% select(BIC)%>% slice(7))
bic65 <- as.numeric(fit %>% select(BIC)%>% slice(8))
awe22 <- as.numeric(fit %>% select(AWE)%>% slice(1))
awe23 <- as.numeric(fit %>% select(AWE)%>% slice(2))
awe24 <- as.numeric(fit %>% select(AWE)%>% slice(3))
awe25 <- as.numeric(fit %>% select(AWE)%>% slice(4))
awe62 <- as.numeric(fit %>% select(AWE)%>% slice(5))
awe63 <- as.numeric(fit %>% select(AWE)%>% slice(6))
awe64 <- as.numeric(fit %>% select(AWE)%>% slice(7))
awe65 <- as.numeric(fit %>% select(AWE)%>% slice(8))
e22 <- as.numeric(fit %>% select(Entropy)%>% slice(1))
e23 <- as.numeric(fit %>% select(Entropy)%>% slice(2))
e24 <- as.numeric(fit %>% select(Entropy)%>% slice(3))
e25 <- as.numeric(fit %>% select(Entropy)%>% slice(4))
e62 <- as.numeric(fit %>% select(Entropy)%>% slice(5))
e63 <- as.numeric(fit %>% select(Entropy)%>% slice(6))
e64 <- as.numeric(fit %>% select(Entropy)%>% slice(7))
e65 <- as.numeric(fit %>% select(Entropy)%>% slice(8))

## Plot best model
`Number of Profiles` <- as.factor(c(2,3,4,5,6,7,8,9))
Profiles <- recode(`Number of Profiles`,"2"="CIRP 2","3"="CIRP 3","4"="CIRP 4","5"="CIRP 5",
                              "6"="CIDP 2","7"="CIDP 3","8"="CIDP 4","9"="CIDP 5")
AIC <- c(aic22,aic23,aic24,aic25,aic62,aic63,aic64,aic65)
BIC <- c(bic22,bic23,bic24,bic25,bic62,bic63,bic64,bic65)
AWE <- c(awe22,awe23,awe24,awe25,awe62,awe63,awe64,awe65)
Entropy <- c(e22,e23,e24,e25,e62,e63,e64,e65)
Model_metrics <- data.frame(Profiles,AIC,BIC,AWE,Entropy)

scl = 35000
figure_2 <- Model_metrics %>%
  ggplot(aes(x = Profiles)) + 
  geom_line(aes(y = AIC, colour = "AIC"), size = 1.0, group=1) +
  geom_line(aes(y = BIC, colour = "BIC"), size = 1.0, group=1) +
  geom_line(aes(y = AWE, colour = "AWE"), size = 1.0, group=1) +
  geom_line(aes(y = Entropy*scl, colour = "Entropy"), size = 1.0,group=1) +
  scale_y_continuous(sec.axis = sec_axis(~. /scl , name = "Entropy")) +
  labs(colour = "Index", y = "AIC / BIC") + 
  theme(legend.position="bottom") + 
  scale_colour_manual(values = c(
    AIC = "royalblue4", BIC = "#E69F00", AWE ="#56B4E9", Entropy ="maroon"))

Model_metrics%>% 
  group_by(Profiles)%>%
  mutate(Total = (sum(AIC,BIC,AWE))/3)%>%
  arrange(Total)

ggsave("Figure_2.tiff", figure_2, width = 10, height = 6, dpi = 300)
```

## CIRP with 4 classes
```{r, echo=FALSE, warning=FALSE, message=FALSE}
write_sav(get_data(CIRP_4), "LPA_with_participant_class.sav")
#Create standardized values in SPSS because R gives you wrong results
data2 <- read_sav("LPA_with_participant_class.sav")

data2%>%
  group_by(Class)%>%
  count(Class)%>%
  mutate(Perc = (n/571)*100)

LPA_data <- data2%>%
  select(Class, ZProteus_1, ZProteus_2, ZProteus_3, ZProteus_4, ZProteus_5, ZProteus_6)%>%
  group_by(Class)%>%
  summarise(ZProteus_1=mean(ZProteus_1),ZProteus_2=mean(ZProteus_2),
            ZProteus_3=mean(ZProteus_3),
            ZProteus_4=mean(ZProteus_4),ZProteus_5=mean(ZProteus_5),ZProteus_6=mean(ZProteus_6))%>%
  na.omit()
LPA_data

LPA_data1 <- LPA_data%>%
  pivot_longer(cols=c(ZProteus_1,ZProteus_2,ZProteus_3,ZProteus_4,ZProteus_5,ZProteus_6),
               names_to = "Model_Indicators", values_to = "Z_Scores")

LPA_data1$Class <- as.factor(LPA_data1$Class)
levels(LPA_data1$Class) <- c("class_1","class_2","class_3", "class_4")

Figure_3 <- LPA_data1%>%
  ggplot(aes(x = Model_Indicators, y = Z_Scores, group = Class, color=Class)) + 
  geom_point(size = 1.8) + geom_line(size = 1.0) +
  labs(x= "", y = "Z scores", color = "Profiles") + 
  theme(axis.title.x = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1.0),
        axis.title.y = element_text(face = "bold"),
        legend.title = element_text(face = "bold"), 
        legend.position="bottom") +
  scale_y_continuous(breaks=seq(-1.0, 2.0, by = 0.5)) +
  scale_x_discrete(labels=c("ZProteus_1"="Proteus_1","ZProteus_2"="Proteus_2",
                            "ZProteus_3"="Proteus_3", 
                            "ZProteus_4"="Proteus_4",
                            "ZProteus_5"="Proteus_5",
                            "ZProteus_6"="Proteus_6")) +
  scale_colour_manual(values = c(
    Class_1 = "royalblue4", `Class_2` = "#E69F00", `Class_3` ="#56B4E9", 'Class_4' ='#e600cf'))

c_data$Class <- data2$Class
#write_sav(c_data, "c_data.sav")
Figure_3
ggsave("Figure_3.tiff", Figure_3, width = 10, height = 6, dpi = 300)
```


